Penalized linear mixed models for structured genetic data

### 2021-02-25 main.tex
Sec 3.8

We generated three varieties of genotype matrices $\bG$, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations \citep{Ochoa2016}, (2) simulated SNP data from independent subpopulations \citep{Ochoa2016}, and (3) real genome-wide SNP data from known Caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}. We refer to these as 1D Linear Admixture, Independent Subpopulations, and Empirical, respectively. For all data types, there are $p = 1000$ total SNPs, 50 of which have causal effects. For the simulated 1D Linear Admixture and Independent Subpopulations data $n = 200$ unless otherwise specified, and $n = 588$ for the Empirical data.

@@

Although cross-validation (CV) and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons. Because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach such as BIC would not lend itself to comparable results across methods. Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples is an open question \citep{roberts2017cross}. \anna{We include a more detailed discussion of tuning parameter selection methods for penalized LMMs in Section \ref{sec:discussion}.}

%------------------------------%
\subsection{\anna{Genomewide analysis}}
%------------------------------%

@@

%------------------------------%
\subsection{\anna{Genome-wide analysis case study}}
%------------------------------%

@@

\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/genomewide_pc1_corr.png}
\caption{Pathwise absolute correlation of selected, non-causal SNPs with the first principal component of the data in Empirical genome-wide analysis. Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular $\lambda$ value, with shaded region indicating $\pm$ 1 SE. Single replication case study with $\eta = 0.5, \xi = 0.8$. \anna{get rid of 'Method' legend title.}}
\label{fig:gw_pc1_corr}
\end{figure}

>>>>

We generated three varieties of genotype matrices $\bG$, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations \citep{Ochoa2016}, (2) simulated SNP data from independent subpopulations \citep{Ochoa2016}, and (3) \anna{a subset of} real genome-wide SNP data from known Caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}. We refer to these as 1D Linear Admixture, Independent Subpopulations, and Empirical, respectively. For all data types, there are $p = 1000$ total SNPs, 50 of which have causal effects. For the simulated 1D Linear Admixture and Independent Subpopulations data $n = 200$ unless otherwise specified, and $n = 588$ for the Empirical data.

@@

Although cross-validation (CV) and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons. Because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach such as BIC would not lend itself to comparable results across methods. Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples remains an open question \citep{roberts2017cross}. \anna{We include a more detailed discussion of tuning parameter selection methods for penalized LMMs in Section \ref{sec:discussion}.}

%------------------------------%
\subsection{\anna{Genome-wide analysis}}
%------------------------------%

To illustrate the statistical properties our simulations aim to elucidate on a more realistic scale, we present an analysis of the full Empirical genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives} with a synthetic phenotype as a pseudo-real data example or case study. We generate the synthetic phenotype as in \eqref{eqn:sim} using 50,000 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and environmental effects ($\xi = 0.8$).

@@

%------------------------------%
\subsection{\anna{LMM-lasso outperforms lasso when analyzing genome-wide data}}
%------------------------------%

@@

\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/genomewide_pc1_corr.png}
\caption{Pathwise absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the data in Empirical genome-wide analysis. Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular $\lambda$ value, with shaded region indicating $\pm$ 1 SE. Single replication case study with $\eta = 0.5, \xi = 0.8$. \anna{get rid of 'Method' legend title.}}
\label{fig:gw_pc1_corr}
\end{figure}
